---
title: "Movielens_Project"
author: "Ocansey Djabaku Kevin"
date: '2022-11-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CAPSTONE PROJECT

```{R Loading files and libraries}
##########################################################
# Create edx set, validation set (final hold-out test set)
##########################################################

# Note: this process could take a couple of minutes

if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")
if(!require(janitor)) install.packages("janitor", repos = "http://cran.us.r-project.org")
if(!require(recosystem)) install.packages("recosystem", repos = "http://cran.us.r-project.org")


library(tidyverse)
library(caret)
library(data.table)
library(janitor)

# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("https://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")

# if using R 3.6 or earlier:
# movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
#                                            title = as.character(title),
#                                            genres = as.character(genres))

# if using R 4.0 or later:
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(movieId),
                                           title = as.character(title),
                                           genres = as.character(genres))


movielens <- left_join(ratings, movies, by = "movieId")

# Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding") # if using R 3.5 or earlier, use `set.seed(1)`
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% 
  semi_join(edx, by = "movieId") %>%
  semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

rm(dl, ratings, movies, test_index, temp, movielens, removed)
```

### Overview and Summary

```{r Summary of Edx }
head(edx)

dim(edx)

edx %>%
  summarize(n_users = n_distinct(userId),
            n_movies = n_distinct(movieId))
```

## VISUALIZATION

```{r Top Movies}
edx %>%
  group_by(title) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  top_n(15, count) %>%
  ggplot(aes(count, reorder(title, count))) +
  geom_bar(color = "black", fill= '#00abff',  stat = "identity") +
  xlab("Count") +
  ylab("Movies")
```

```{r Ratings per users }
    edx %>% group_by(userId) %>%
      summarize(count = n()) %>%
      ggplot(aes(count)) +
      geom_histogram(color = "black", fill = "#00abff", bins = 40) +
      xlab("Ratings") +
      ylab("Users") +
      scale_x_log10() 
```

#### Dividing edx into training and testing datasets

```{r Training and Test Sets }
index_edx <-  createDataPartition(y = edx$rating, times = 1, p = 0.8, list =  FALSE) 
train_set <- edx[index_edx,]
test_set <- edx[-index_edx,]

test_set <- test_set %>%
  semi_join(train_set, by = "movieId") %>%
  semi_join(train_set, by = "userId")

```

#### RMSE FUNCTION

```{r Rmse Function}
RMSE <- function(true_ratings, predicted_ratings){
  sqrt(mean((true_ratings - predicted_ratings)^2))
}
```

## MACHINE LEARNING

### Simple Model

```{r Simple Model using average ratings}
# We use the avg of the ratings to build a simple predictive model.
mu_ratings <- mean(train_set$rating)
mu_ratings

# Our simple model looks like this mathematically  =>  Yu,i = μ_hat +  εu,i
# where Y represents our predicted rating, u_hat our avg rating for all movies and epsolon represnts any minor error 

simple_model_rmse <- RMSE(test_set$rating, mu_ratings)

# Meanwhile i would create a table to keep track of all the rmse results as we improve the model
rmse_results <- tibble(method = "Simple Model using Average", RMSE = simple_model_rmse)
```

### Simple Model + Movie Effect

```{r Simple Model + Movie bias/effect}
# 2. Movie effect model + simple model.   
mu <- mean(train_set$rating)
movie_avgs <- train_set %>%
  group_by(movieId) %>%
  summarize(b_i = mean(rating - mu))

# Now adding b_i into our first simple model gives us something like this  yu,i = μ_hat + b_hati
predicted_ratings <- mu + test_set %>%
  left_join(movie_avgs, by='movieId') %>%
  pull(b_i)
movie_effect <- RMSE(predicted_ratings, test_set$rating)


rmse_results <- add_row(.data = rmse_results, method= 'Simple Model + Movie effects', RMSE =  movie_effect )
```

#### Movie Bias plot

```{r Movie Effect Plot and updated Rmse Table }
movie_avgs %>% ggplot(aes(b_i)) +
  geom_histogram(color = "black", fill = "#00abff", bins = 10) +
  xlab("Movie Bias") +
  ylab("Count") 

rmse_results
```

### Simple Model + Movie effect + User effect

```{r Simple Model + Movie effect + User effect}
# User Effect Model
user_avgs <- train_set %>%
  left_join(movie_avgs, by='movieId') %>%
  group_by(userId) %>%
  summarize(b_u = mean(rating - mu - b_i))

predicted_ratings <- test_set %>%
  left_join(movie_avgs, by='movieId') %>%
  left_join(user_avgs, by='userId') %>%
  mutate(pred = mu + b_i + b_u) %>%
  pull(pred)
user_effect <- RMSE(predicted_ratings, test_set$rating)


rmse_results <- add_row(.data = rmse_results, method = 'Simple Model + Movie Effect + User effects', RMSE = user_effect )

# Now We  need to get an Rmse lower than 0.86
```

#### User Bias plot

```{r User bias plot and Update }
train_set %>%
  group_by(userId) %>%
  summarize(b_u = mean(rating)) %>%
  filter(n()>=100) %>%
  ggplot(aes(b_u)) +
  geom_histogram(bins = 30, fill = "#00abff")

rmse_results
```

## REGULARIZATION

```{r Regularization}
# Regularization
# This is important because we need to make sure that movies with little as 30 or 100 ratings have similar weight as movies with 1000 plus ratings

lambdas <- seq(0, 10, 0.25)
rmses <- sapply(lambdas, function(x){
  b_i <- train_set %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n()+x))
  b_u <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    group_by(userId) %>%
    summarize(b_u = sum(rating - b_i - mu)/(n()+x))
  predicted_ratings <- test_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    mutate(pred = mu + b_i + b_u) %>%
    pull(pred)
  return(RMSE(predicted_ratings, test_set$rating))
})

lambda_rmse <-  lambdas[which.min(rmses)]
qplot(lambdas, rmses)
# lambda of 4.75 provides the lowest RMSE error

rmse_results <- add_row(.data = rmse_results, method= 'Regularized Movies + User effects + ', RMSE = min(rmses))

rmse_results
```

## MATRIX FACTORIZATION WITH RECOSYSTEM

### This will take some time, but i promise its worth it

```{r recosystem factorization}
library(recosystem)
set.seed(1, sample.kind="Rounding")
train_reco <- with(train_set, data_memory(user_index = userId, item_index = movieId, rating = rating))
test_reco <- with(test_set, data_memory(user_index = userId, item_index = movieId, rating = rating))
r <- Reco()

tuned_reco <- r$tune(train_reco, opts = list(dim = c(20, 30),
                                            costp_l2 = c(0.01, 0.1),
                                            costq_l2 = c(0.01, 0.1),
                                            lrate = c(0.01, 0.1),
                                            nthread = 4,
                                            niter = 10))

r$train(train_reco, opts = c(tuned_reco$min, nthread = 4, niter = 30))
results_reco <- r$predict(test_reco, out_memory())


factorization_rmse <- RMSE(results_reco, test_set$rating)

rmse_results <- add_row(.data = rmse_results, method= 'Factorization using Reco library', RMSE = factorization_rmse)
# our Rmse now is below 0.86 at 0.79
```

## FINAL TESTING WITH OUR VALIDATION DATA SET

```{r Train with edx data set and test with validation rating}

set.seed(1, sample.kind="Rounding")
edx_reco <- with(edx, data_memory(user_index = userId, item_index = movieId, rating = rating))
validation_reco <- with(validation, data_memory(user_index = userId, item_index = movieId, rating = rating))
r <- Reco()

tuned_reco <- r$tune(edx_reco, opts = list(dim = c(20, 30),
                                          costp_l2 = c(0.01, 0.1),
                                          costq_l2 = c(0.01, 0.1),
                                          lrate = c(0.01, 0.1),
                                          nthread = 4,
                                          niter = 10))

r$train(edx_reco, opts = c(tuned_reco$min, nthread = 4, niter = 30))

final_reco <- r$predict(validation_reco, out_memory())
final_rmse <-  RMSE(final_reco,validation$rating)
rmse_results <- add_row(.data = rmse_results, method= 'Final Rmse', RMSE = final_rmse)

```

### Our RMSE Table

```{r RMSE TABLE}
rmse_results
```
